### 2016-12-09 related-work.tex
continuing my pass, cleaning up the .bib files

Naor and Yogev~\cite{C:NaoYog15} recently considered the problem of
adversarial soundness for Bloom filters and, indeed, their work
served as inspiration for our own. Our work extends their work
significantly in several directions. For starters, we consider
abstract data structures (rather than only the case of
set-membership data structures), and consider privacy in addition to
correctness. Even with respect to the specific case of correctness
for set-membership data structures, our work offers several
advantages to the Naor-Yogev treatment:
\begin{itemize}
\item We explicitly distinguish between public and private
    portions of a data structure, something they leave
    ambiguous.\footnote{In fact, it isn't clear that their
formalization actually allows for standard Bloom filters,
    although fixing this ambiguity is straightforward once
    recognized.} This allows us, for example, to distinguish
    between the hash-function keys and the bit array when
    discussing classical Bloom filters.
\item The Naor-Yogev definition of soundness allows the
    adversary to make several set-membership queries, some of
    which may produce incorrect results, and the attacker then
    succeeds if it can output a \emph{fresh} query which leads
    to an error. We propose instead what we find to be a more
    natural definition in which the attacker succeeds if it can
    cause a certain number of incorrect results.
\item Naor and Yogev analyze a new Bloom-filter construction of
    their own design. In contrast, we are interested in
    analyzing practical constructions that have been proposed in
    prior work to understand their security. In particular, we
    analyze Bloom filters in the random-oracle model, something
    not considered by Naor and Yogev.
\end{itemize}

@@

\jnote{Rather random collection of papers using Bloom filters and
variants. I removed it for now, since it's not clear that they have
any particular relevance to us.}
%As previously mentioned, Bloom
%filters and their relatives are some of the most widely used data
%structures supporting set-membership queries. As examples, Hbase,
%the open-source implementation of Google's BigTable storage
%system~\cite{chang2008bigtable},
%%a Hadoop-based NoSql database designed to handle large datasets,
%includes an implementation of Bloom filters and counting Bloom
%filters, and the Squid proxy~\cite{fan2000summary} uses a Bloom
%filter as a ``summary'' of the set of URLs in its cache in order to
%improve latency for web-object retrieval. Reynolds and
%Vahdat~\cite{reynolds2003efficient} proposed an efficient
%distributed search engine that can be used to search for files
%containing a particular keyword. Their search engine maps the
%keywords of each file into a Bloom filter; a look-up of the keyword
%in the Bloom filter tells whether the node has files containing that
%keyword or not. Stochastic Fair Blue~\cite{feng2001stochastic} uses
%counting Bloom filter to manage non-responsive TCP traffic.

>>>>

Naor and Yogev~\cite{C:NaoYog15} recently considered the problem of
adversarial soundness for Bloom filters and, indeed, their work
served as inspiration for our own. Our work extends their work
significantly in several directions. For starters, we consider
abstract data structures (rather than only the case of
set-membership data structures), and consider privacy in addition to
correctness. Even with respect to the specific case of correctness
for set-membership data structures, our work offers several
advantages to the Naor-Yogev treatment:
\begin{itemize}
\item We explicitly distinguish between public and private
    portions of a data structure, something they leave
    ambiguous.\footnote{In fact, it isn't clear that their
formalization allows for standard Bloom filters, though fixing
    this ambiguity is straightforward once recognized.} This
    allows us, for example, to distinguish between the
    hash-function keys and the bit array when discussing
    classical Bloom filters.
\item The Naor-Yogev definition of soundness allows the
    adversary to make several set-membership queries, some of
    which may produce incorrect results, and the attacker then
    succeeds if it can output a \emph{fresh} query which leads
    to an error. We propose instead what we find to be a more
    natural definition in which the attacker succeeds if it can
    cause a certain number of incorrect results.
\item Naor and Yogev analyze a new Bloom-filter construction of
    their own design. In contrast, we are interested in
    analyzing practical constructions that have been proposed in
    prior work to understand their security. In particular, we
    analyze Bloom filters in the random-oracle model, something
    not considered by Naor and Yogev.
\end{itemize}

@@

\jnote{Rather random collection of papers using Bloom filters and
variants. I removed it for now, since it's not clear that they have
any particular relevance to us. I kept only the refs that seemed
directly relevant.}
%As previously mentioned, Bloom
%filters and their relatives are some of the most widely used data
%structures supporting set-membership queries. As examples, Hbase,
%the open-source implementation of Google's BigTable storage
%system~\cite{chang2008bigtable},
%%a Hadoop-based NoSql database designed to handle large datasets,
%includes an implementation of Bloom filters and counting Bloom
%filters, and the Squid proxy~\cite{fan2000summary} uses a Bloom
%filter as a ``summary'' of the set of URLs in its cache in order to
%improve latency for web-object retrieval. Reynolds and
%Vahdat~\cite{reynolds2003efficient} proposed an efficient
%distributed search engine that can be used to search for files
%containing a particular keyword. Their search engine maps the
%keywords of each file into a Bloom filter; a look-up of the keyword
%in the Bloom filter tells whether the node has files containing that
%keyword or not. Stochastic Fair Blue~\cite{feng2001stochastic} uses
%counting Bloom filter to manage non-responsive TCP traffic.

### 2016-12-09 results.tex
continuing my pass, cleaning up the .bib files

\section{Case Studies}
Here we explore various set-membership data structures that have been proposed
in the literature with respect to our definitions.

@@

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets F_{K}(\langle j,x \rangle) $\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1
}
\caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for classic Bloom
  filter $\setprim_{\mathrm{Bloom}}$ in the ROM for a hash function $H\colon \mathbb{N} \times
  \elts \to [m]$. {\bf Right:} $\Rep$ and $\Qry$
  algorithms for the domain-separated PRF data structure
  $\setprim_{\mathrm{ds}}$, using function family $F \colon
  \calK \times \elts \to [m]$.  Both constructions have query set is $\calQ=\{q_x\colon
  \univ \to \bits\}$ where $q_x(S)=1 \Leftrightarrow x \in S$.  }
\label{fig:bf-and-ds}
\end{figure}

\subsection{Classical Bloom Filter in the Random-Oracle Model}
Fix $n,k,m \geq 0$, and fix a set $\elts$.  Then the classical Bloom
filter set-membership data structure $\setprim_{\mathrm{Bloom}}=
(\Rep, \Qry)$ over~$\elts$ is defined as in
Figure~\ref{fig:bf-and-ds} (left side).  We
define~$\setprim_{\mathrm{Bloom}}$ in the random-oracle model
(ROM), as the traditional analysis of Bloom filter correctness (as a
function of $n,k,m$) is done in this model.\footnote{The standard
model version would have $\Rep$ choose $H_1,H_2,\ldots,H_k$ from a
hash function family, compute $h_j$ using these, and placing their
description in $\pubaux$.}

\heading{Correctness. } The following theorem proves that the Bloom
filter meets our correctness notion in the ROM.  Again, we stress
that this correctness is with respect to adaptive adversaries,
aiming to find $r \geq 1$ errors (in this case, false-positives).

\begin{theorem}\label{thm3}\label{thm:bf-correctness}
Fix $k,m,n,r>0$ and a set~$\elts$, and let
$\setprim_{\mathrm{Bloom}}= (\Rep, \Qry)$ be the classical Bloom
filter (set-membership) data structure over~$\elts$. For any
adversary~$A$ that makes a total of~$q_T$ queries to the $\Test$
oracle, and $q_R$ queries to the RO, and has time-complexity~$O(t)$,
\[
\AdvCorrect{\setprim_{\mathrm{Bloom}},r}{A} \leq  {\dbinom{q_T + q_H-1}{r-1}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
\end{theorem}
We note that Kirsch and Mitzenmacher~\cite{kirsch2006less} show
that the ($r=1$, zero-query) error-probability for~$\setprim_\mathrm{Bloom}$ is
always within $O(1/n)$ of $(1-e^{-kn/m})^k$, although the asymptotic
rate of convergence is faster than~$1/n$.  The multiplicative (``hybrid'') term arises as a result
of moving from the adaptive to the zero-query setting.
Finally, $\binom{q_T+q_H-1}{r-1} \leq (q_T+q_H)^r$ (with reasonable tightness when~$r$ is small) in
which case the final term in the bound behaves as
$((q_T+q_H) (1-e^{-kn/m})^k + O(q/n) )^r$; loosely, this is the bound one expects for
trying to find a single correctness error (in this case, a false positive) in each of~$r$ independent``rounds'', each round
consisting of~$(q_T+q_H)$ attempts.
\input{BF_correctness}

>>>>

\section{Case Studies}\label{sec:case_studies}
Here we explore various set-membership data structures that have been proposed
in the literature with respect to our definitions.

@@

\medskip
\algorithmv{$\Qry(\pubaux,\privaux,q_x)$}\\
$M\gets\pubaux$, $K \gets \privaux$\\
for $j \in \{1,2,\ldots,k\}$\\
\nudge $h_j \gets F_{K}(\langle j,x \rangle) $\\
\nudge if $M[h_j] \neq 1$ then Return 0\\
Return 1 } \caption{{\bf Left:} $\Rep$ and $\Qry$ algorithms for
classical Bloom
  filter $\setprim_{\mathrm{Bloom}}$ in the ROM for a hash function $H\colon \mathbb{N} \times
  \elts \to [m]$, where $|S| \leq n$. {\bf Right:} $\Rep$ and $\Qry$
  algorithms for the domain-separated PRF data structure
  $\setprim_{\mathrm{ds}}$, using function family $F \colon
  \calK \times \elts \to [m]$.  Both constructions have query set is $\calQ=\{q_x\colon
  \univ \to \bits\}$ where $q_x(S)=1 \Leftrightarrow x \in S$.  }
\label{fig:bf-and-ds}
\end{figure}

\subsection{Classical Bloom Filter in the Random-Oracle Model}
Fix $k,m,n \geq 0$ and a universe $\elts$.  Then the classical Bloom
filter set-membership data structure $\setprim_{\mathrm{Bloom}}=
(\Rep, \Qry)$ over~$\elts$ is defined as in
Figure~\ref{fig:bf-and-ds} (left side) for sets $S \subseteq \elts$
of size at most $|S|\leq n$. We define~$\setprim_{\mathrm{Bloom}}$
in the random-oracle model (ROM), as the traditional analysis of
correctness (as a function of $n,k,m$) is done in this
model.\footnote{The standard-model version would have $\Rep$ choose
$H_1,\ldots,H_k$ from some function family, compute $h_j$ using
these, and place their description in $\pubaux$.}

\heading{Correctness. } We prove that the Bloom filter meets our
correctness notion in the ROM.  Again, we stress that this
correctness is with respect to adaptive adversaries, aiming to find
$r \geq 1$ errors (in this case, false-positives).

\begin{theorem}\label{thm3}\label{thm:bf-correctness}
Fix $k,m,n,r>0$ and~$\elts$, and let $\setprim_{\mathrm{Bloom}}=
(\Rep, \Qry)$ be the classical Bloom filter (set-membership) data
structure over~$\elts$. For any adversary~$A$ that makes a total
of~$q_T$ queries to the $\Test$ oracle, $q_R$ queries to the random
oracle, and has time-complexity~$O(t)$,
\[
\AdvCorrect{\setprim_{\mathrm{Bloom}},r}{A} \leq  {\dbinom{q_T + q_H-1}{r-1}} \left( (1-e^{-kn/m})^k + O(1/n) \right)^r\,.
\]
\end{theorem}
Note that Kirsch and Mitzenmacher~\cite{kirsch2006less} show that
the ($r=1$, zero-query) error-probability
for~$\setprim_\mathrm{Bloom}$ is %always within $O(1/n)$ of
$(1-e^{-kn/m})^k+O(1/n)$. %, although the asymptotic rate of
%convergence is faster than~$1/n$.
The multiplicative %(``hybrid'')
term arises as a result of moving from the zero-query to the
adaptive setting. Also note that $\binom{q_T+q_H-1}{r-1} \leq
(q_T+q_H)^r$ (with reasonable tightness when~$r$ is small), in which
case the bound becomes
\[\left((q_T+q_H) \cdot (1-e^{-kn/m})^k + O((q_T+q_H)/n) \right)^r;\]
loosely, and ignoring the low-order term, this is the bound one
expects if we view the attacker as trying to find an error in each
of~$r$ independent``rounds,'' each round consisting of~$q_T+q_H$
attempts. \jnote{But if the attacker makes $q=q_T+q_H$ queries
overall, isn't it only making $q/r$ queries per round?}
%JK
\input{BF_correctness}

### 2016-12-09 syntax.tex
intro
continuing my pass, cleaning up the .bib files

\newcommand{\qry}{\mathsf{qry}}
\section{Preliminaries}

When~$\univ$ is a set, we write $x \getsr \univ$ to denote sampling
an element uniformly from~$\univ$ and assigning it to~$x$. We extend
this notation to randomized algorithms, so $x \getsr A(y)$ denotes
that (randomized) algorithm $A$ is run on input~$y$, and the output
of the algorithm is assigned to~$x$. Algorithms may be provided
black-box access to one or more oracles, which we write as
superscripts, e.g., $F^{O_1,O_2,\ldots}$.  All algorithms (and, in
particular, all adversaries) are stateful by default. We write
$\langle x,y,\ldots \rangle$ for the encoding of values $x,y,\ldots$
as a single bit-string, for some implicit and invertible encoding
method. When $x$ is a string, we write $x[i]$ for the $i$-th element
in that string, and $x[i:j]$ for the substring $x[i]x[i+1]\cdots
x[j]$.  \tsnote{need subset of size~$r$ notation}

@@

%Informally, the representation should serve as a good
%approximation of the multiset, at least with respect to the allowed queries.
\begin{definition}[Data structure] \rm
Fix sets~$\univ, \Gamma$, and $\calQ = \{\qry \colon \univ \to
\Gamma\}$. A \emph{data structure} (supporting queries $\calQ$ over
universe $\univ$) is a tuple $\setprim=(\Rep, \Qry)$ where:
\begin{itemize}
\item $\Rep \colon \univ \to \bits^* \times \bits^*$ is a
    randomized \emph{representation algorithm}, taking as input
    an element $S \in \univ$,
    and outputting %a \emph{representation} $M \in \Sigma^*$,
    \emph{public representation data} $\pubaux \in \bits^*$ and
    \emph{private representation data} $\privaux \in \bits^*$.
    We write $(\pubaux,\privaux) \getsr \Rep(S)$ as shorthand
    for this execution.
\item $\Qry\colon \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $\pubaux,\privaux \in \bits^*$
    and a query $\qry \in \calQ$, and outputting an
    \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(\pubaux,\privaux,\qry)$ for this execution. \hfill\dqed
\end{itemize}
\end{definition}

\todo{Insert discussion of syntax, if needed.  If we think it is
  helpful to discuss the size of the representation, we can declare it
as $|\pubaux|+|\privaux|$ and comment that if one wants to compare to
classical size bounds, a more fine-grained separation of the output of
$\Rep$ would allow this.}

\begin{definition}[Error probability and rate] \rm
For a specific data structure $\setprim=(\Rep, \Qry)$ supporting
queries $\calQ$ over universe $\univ$, let
$\mathrm{Err}(S,\qry,\pubaux,\privaux)$ be  true iff
$\Qry(\pubaux,\privaux, \qry) \neq \qry(S)$.  Then we define for
$\setprim$ the quantities
\begin{align*}
&\max_{S,\qry}\Prob{(\pubaux,\privaux) \getsr \Rep(S) :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1 } &\\
&\max_{S}\Prob{(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1} &
\end{align*}
which are the \emph{error probability} and \emph{error rate}, respectively.
%
%We say it has \emph{error rate} $\tilde{\epsilon}$ if
%\[{\textstyle \max_{S}\Pr[(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
%\mathrm{Err}(S,\qry,\pubaux,\privaux)=1]} \leq \tilde{\epsilon}. \]
\hfill\dqed
\end{definition}
\todo{Words :)  Can compare the (normal) error probability to
  the adversarial soundness probability.  Perhaps pull back some of
  the previous example for building intuition about the difference
  between the two.}

@@

The above definition can be lifted to the random-oracle model by
giving $\Rep, \Qry$ access to a random oracle. In that case, the
probabilities are also taken over choice of the random oracle. We
remark that, while our error probability and error rate do not
distinguish among ``types'' of errors, they also do not preclude
instantiations (like the classical Bloom filter) that admit only one
type of error, e.g., false positives.

@@

%\heading{Data structures for set-membership queries.}
An important special case is a data structure that supports set-membership
queries. \jnote{I changed this from multiplicity to membership.}

\begin{definition} \rm
A \emph{set-membership data structure over $\elts$} is a data
structure with $\univ \subseteq 2^\elts$, $\Gamma=\{0,1\}$, and
$\calQ=\{\qry_x\}_{x \in \univ}$, where $\qry_x(S)$ is true iff $x
\in S$. \hfill\dqed
\end{definition}

@@

>>>>

\section{Preliminaries}

When~$\univ$ is a set, we write $x \getsr \univ$ to denote sampling
a uniform element from~$\univ$ and assigning it to~$x$. We extend
this notation to randomized algorithms, so $x \getsr A(y)$ denotes
that (randomized) algorithm $A$ is run on input~$y$, and the output
of the algorithm is assigned to~$x$. Algorithms may be provided
black-box access to one or more oracles, which we write as
superscripts, e.g., $F^{O_1,O_2,\ldots}$.  All algorithms (and, in
particular, all adversaries) are stateful by default. We write
$\langle x,y,\ldots \rangle$ for the encoding of values $x,y,\ldots$
as a single bit-string, for some implicit and invertible encoding
method. When $x$ is a string, we write $x[i]$ for the $i$-th element
in that string, and $x[i:j]$ for the substring $x[i]x[i+1]\cdots
x[j]$.  %\tsnote{need subset of size~$r$ notation}

@@

%Informally, the representation should serve as a good
%approximation of the multiset, at least with respect to the allowed queries.
\begin{definition}[Data structure] \rm
Fix sets~$\univ, \Gamma$, and $\calQ = \{\qry \colon \univ \to
\Gamma\}$. A \emph{data structure} (supporting queries $\calQ$ over
universe $\univ$) is a tuple $\setprim=(\Rep, \Qry)$ where:
\begin{itemize}
\item $\Rep \colon \univ \to \bits^* \times \bits^*$ is a
    randomized \emph{representation algorithm}, taking as input
    $S \in \univ$,
    and outputting %a \emph{representation} $M \in \Sigma^*$,
    \emph{public representation data} $\pubaux \in \bits^*$ and
    \emph{private representation data} $\privaux \in \bits^*$.
    We write $(\pubaux,\privaux) \getsr \Rep(S)$ as shorthand
    for this execution.
\item $\Qry\colon \bits^* \times \bits^* \times \calQ \to
    \Gamma$ is a deterministic \emph{query-evaluation}
    algorithm, taking as input $\pubaux,\privaux \in \bits^*$
    and a query $\qry \in \calQ$, and outputting an
    \emph{answer} $a \in \Gamma$.  We write $a \gets
    \Qry(\pubaux,\privaux,\qry)$ for this execution. \hfill\dqed
\end{itemize}
\end{definition}

\begin{definition}[Error probability and rate] \rm
For a specific data structure $\setprim=(\Rep, \Qry)$ supporting
queries $\calQ$ over universe $\univ$, let
$\mathrm{Err}(S,\qry,\pubaux,\privaux)$ be  true iff
$\Qry(\pubaux,\privaux, \qry) \neq \qry(S)$.  Then we define
\begin{align*}
&\max_{S,\qry}\Prob{(\pubaux,\privaux) \getsr \Rep(S) :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1 } &\\
&\max_{S}\Prob{(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
\mathrm{Err}(S,\qry,\pubaux,\privaux)=1} &
\end{align*}
to be, respectively, the \emph{error probability} and \emph{error
rate} of~$\Pi$.
%
%We say it has \emph{error rate} $\tilde{\epsilon}$ if
%\[{\textstyle \max_{S}\Pr[(\pubaux,\privaux) \getsr \Rep(S);\, \qry \getsr \calQ :
%\mathrm{Err}(S,\qry,\pubaux,\privaux)=1]} \leq \tilde{\epsilon}. \]
\hfill\dqed
\end{definition}

@@

The above definitions can be lifted to the random-oracle model by
giving $\Rep, \Qry$ access to a random oracle. In that case, all
probabilities are also taken over choice of the random oracle. We
remark that, while our error probability and error rate do not
distinguish among different ``types'' of errors, they also do not
preclude instantiations (like the classical Bloom filter) that admit
only one type of error, e.g., false positives.

@@

%\heading{Data structures for set-membership queries.}
An important special case is a data structure for set-membership
queries.
\begin{definition} \rm
A \emph{set-membership data structure over $\elts$} is a data
structure with $\univ \subseteq 2^\elts$, $\Gamma=\{0,1\}$, and
$\calQ=\{\qry_x\}_{x \in \univ}$ where $\qry_x(S)$ is true iff $x
\in S$. \hfill\dqed
\end{definition}

@@

\heading{The role of $\pubaux$.} One can always ``increase
security'' by pushing all representation data into $\privaux$ and
making $\pubaux$ empty. (This is true in a formal sense for the
definitions given subsequently, but is intuitively true regardless.)
But this misses the point. First, we are interested in analyzing
existing data structures, some of which have a non-empty public
portion. Moreover, for some applications a non-trivial $\pubaux$ is
essential; this will become clear in Section~\ref{sec:case_studies}.
